---
title: "Managing missing values in `ggene`"
slug: "ggene3"
author: "Jean-Pierre Rossi"
date: 2016-09-03T21:13:14-05:00
categories: ["R"]
tags: ["ggene", "genetic", "geostatistics"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggene)
```

Missing values are a big pain in the arse because they are so common in microsatellite datasets that removing them sometimes amounts to loose all the information (individuals or alleles).

`ggene` cannot handle NAs and thus users must remove individuals/alleles with missing data or convert them to something… else!

One common strategy is to replace NAs by the mean value of the corresponding column or by 0 (see the package adegenet). I won’t discuss these options and simply provide a simple solution to users who cannot remove NAs in their data and have to replace them.

We define below two functions `NAtoMean` and `NAto0` which replace NA by the mean of the column (allele) or by 0, respectively.

```{r, message=FALSE, warning=FALSE}
NAtoMean <- function(x){
	x[is.na(x)] <- mean(x, na.rm = TRUE)
	return(x)}

NAto0 <- function(x){ 
	x[is.na(x)] <- 0
	return(x)}
```

Now we create a missing data in the `$tab` of a ggene object:

```{r, message=FALSE, warning=FALSE}
data(crypho)
crypho$tab[1,1]
crypho$tab[1,1] <- NA
crypho$tab[1,1]
```

If we try to compute the variogram with this altered dataset we get an error message:

```{r, error=TRUE, message=T, warning=T}
var <- svariog(crypho, plot=TRUE)
```

Replacing NAs by the mean of the column in the data table is achieved using `apply`: 

```{r, message=FALSE, warning=FALSE}
crypho$tab <- apply(X=crypho$tab, MARGIN=2, FUN=NAtoMean)
crypho$tab[1,1]
```

The `svariog` can now compute the variogram:

```{r, message=FALSE, warning=FALSE}
var <- svariog(crypho, plot=TRUE)
```

`NAtoMean` and `NAto0` can easily be customised.
